---
title: "workstation"
author: "RF"
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(mizerEvolution)
library(mizerHowTo)
library(viridis)
library(tictoc)

```




```{r}
params <- evoParams(no_sp =  5)

sim <- project(params)

plotBiomass(sim)

plotDiet(sim@params, species = "5")

simEvo <- evoProject(params = params, t_max = 200, mutation = 3, trait = "beta")

plotDynamics(simEvo)

plotevoTrait(simEvo, traitID = "beta", returnData = F)

```


### Notes

for creating new traits, the specific cases are not working. for now any trait in species_params can be randomly changed but it doesnt change any calculations around the parameter / trait
need to rewrite the cases to remove all the useless stuff

getSMort to calculate starvation is missing, check mizer or some old files. Maybe mizer does it better anyway

not sure feeding level through time works well, probably useless anyway, need to remove it


## invasions

```{r}

params <- evoParams(no_sp =  5)

simEvo <- evoProject(params = params, t_max = 100, alien = 3, mutation = 0)

```


Adding a dataframe containing the trait range for invasions. For now the df is going to be flying around but it would be good if I could add it to the mizerParams object



```{r}


# using the median from NS to fill df

small_beta <- NS_params@species_params$beta[-c(1,2,4)]
big_beta <- NS_params@species_params$beta[c(1,2,4)]
w_inf_fit <- MASS::fitdistr(NS_params@species_params$w_inf,"lognormal")
beta_fit1 <- MASS::fitdistr(small_beta,"normal") # maybe beta is not the best fit for beta, need a u shaped function, I don't like the logistic either, trying sum of 2 normal
beta_fit2 <- MASS::fitdistr(big_beta,"normal") 
sigma_fit <- MASS::fitdistr(NS_params@species_params$sigma,"normal")
k_vb_fit <- MASS::fitdistr(NS_params@species_params$k_vb,"normal")
ks_fit <- MASS::fitdistr(NS_params@species_params$ks,"normal")
eta_fit <- MASS::fitdistr(NS_params@species_params$w_mat/NS_params@species_params$w_inf,"normal")

# sample new trait values from these fits
fit_winf <- rlnorm(x, w_inf_fit$estimate['meanlog'], w_inf_fit$estimate['sdlog'])

fit_beta1 <- rnorm(x, beta_fit1$estimate['mean'], beta_fit1$estimate['sd'])
fit_beta1[fit_beta1<0] <- NA
fit_beta2 <- rnorm(x, beta_fit2$estimate['mean'], beta_fit2$estimate['sd'])
fit_beta2[fit_beta2<0] <- NA
fit_beta <- c(fit_beta1, fit_beta2)

fit_sigma <- rnorm(x, sigma_fit$estimate['mean'], sigma_fit$estimate['sd'])

fit_kvb <- rnorm(x, k_vb_fit$estimate['mean'], k_vb_fit$estimate['sd'])

fit_ks <- rnorm(x, ks_fit$estimate['mean'], ks_fit$estimate['sd'])

fit_plot <- data.frame(x = fit_ks) #


# they all look fine now
ggplot(fit_plot) +
    geom_density(aes(x =x )) 
    # geom_histogram(aes(x=x))
    # scale_x_continuous(trans = "log10")
    # geom_density(data = test, aes(x = y))


# default dataframe

trait_range <- data.frame("trait" = c("w_inf", "betaS","betaL", "sigma", "k_vb", "ks","eta"),
                          "distribution" = c("lnorm","norm", "norm","norm","norm","norm","norm"),
                          "mean" = c(6.7648459,186.22222,243488.33,1.7166667,0.44408333,6.2893097,0.12071530), # values extracted from NS_params
                          "sd" = c(2.2524833,176.59733,144374.82,0.5742144,0.27293481,2.5737627,0.11480856))


# creating an invader from this

alien_synthesis <- function(trait_range, n = 1){
    
    if(is.null(trait_range)) # using NS_params
    {
        trait_range <- data.frame("trait" = c("w_inf", "betaS","betaL", "sigma", "k_vb", "ks","eta"),
                          "distribution" = c("lnorm","norm", "norm","norm","norm","norm","norm"),
                          "mean" = c(6.7648459,186.22222,243488.33,1.7166667,0.44408333,6.2893097,0.12071530),
                          "sd" = c(2.2524833,176.59733,144374.82,0.5742144,0.27293481,2.5737627,0.11480856))
    }
    
    for(iAlien in 1:n) # for n number of alien
    {
        # for(iRow in dim(trait_range)[1]) # go through each row of the trait df
        # {
        # hard to make it user fool proof so going to focus on default df for now
        
        w_inf <- sigma <- k_vb <- ks <- eta <- beta <- -1 # initialisation for while loop
        
        while(w_inf <0) w_inf <- rlnorm(1, trait_range$mean[1], trait_range$sd[1])
        while(sigma <0)    sigma <- rnorm(1, trait_range$mean[4], trait_range$sd[4])
        while(k_vb <0)    k_vb <- rnorm(1, trait_range$mean[5], trait_range$sd[5])
        while(ks <0)    ks <- rnorm(1, trait_range$mean[6], trait_range$sd[6])
        while(eta <0.027)    eta <- rnorm(1, trait_range$mean[7], trait_range$sd[7]) # eta can get really small in NS_params, just making a threshold at the min value of the ecosystem
        while(beta <0)
        {
            betaS <- rnorm(1, trait_range$mean[2], trait_range$sd[2])
            betaL <- rnorm(1, trait_range$mean[3], trait_range$sd[3])
            beta <- sample(c(betaS,betaL),1)
        }
        w_mat <- w_inf * eta
        
        species_df <- data.frame(
            "w_inf" = w_inf,
            "w_mat" = w_mat,
            "beta" = beta,
            "sigma" = sigma,
            "k_vb" = k_vb,
            "ks" = ks)
    }
    return(species_df)
   }

        
    
    
    trait_range <- data.frame("trait" = c("w_inf", "beta", "sigma", "k_vb", "ks","eta"),
                              "distribution" = c("lnorm","uform", "uform","uform","uform","norm"),
                              "var1" = c(6.7648459,10,0.8,0.1,2.8,0.12071530),
                              "var2" = c(2.2524833,400000,3.2,1,13,0.11480856),row.names = 1)




# tests

x <- 1:1000
y <- rnorm(x)
z <- rlnorm(x)

plot_dat <- data.frame(x = rep(x,2), y=c(y,z), "distrib" = c(rep("norm",1000),rep("logn",1000)))
ggplot(plot_dat) +
    geom_density(aes(x=y)) +
    facet_grid(~distrib)





```





## testing invasions

```{r}

params <- evoParams(updateParams = NS_params)
tic()
sim <- evoProject(params = params, mutation = 0, alien = 5, t_max = 1500)
toc()

# testing TBM

params <- evoParams(no_sp = 9)
sim <- evoProject(params = params, mutation = 0, alien = 5, t_max = 1500)

```

List of errors

Some of your species have an maximum size larger than max_w
Error in get_h_default(params) : Could not calculate h

Warning in max(which(phi > 0)) :
  no non-missing arguments to max; returning -Inf
Error in (ri + 1):2 : result would be too long a vector

Warning in file(file, "wb") :
  cannot open file '/data/home/romainf/mizerEvolution/.Rproj.user/shared/notebooks/D4D53583-romainZone/1/7B36BAD944be3418/c92wx7cuocphw_t/_rs_rdf_839c14aaf46f.rdf': No such file or directory
Error in file(file, "wb") : cannot open the connection

## Diagnostic plots


```{r}

# length of survival per species
survivalLength <- sim@params@species_params$ext - sim@params@species_params$pop # negative values means the species was still alive at the end of the sim
names(survivalLength) <- sim@params@species_params$species
survivalRank <- sort(survivalLength) # this sort put all the surviving species first, then the extinct ones (but in inverted order)
survivalRank <- c(survivalRank[which(survivalRank <0)],survivalRank[rev(which(survivalRank >= 0))]) # now they are in the right order, need to assign weights

# for now, you get 1 if you survived and 0 if you didn't
survivalRank[which(survivalRank >= 0)] <- 0
survivalRank[which(survivalRank < 0)] <- 1
survivalRank <- survivalRank[order(as.numeric(names(survivalRank)))]
    
# doing different order because don't know which one I pref


# % of total community biomass per species
biom <- getBiomass(sim)[dim(sim@n)[1],]
biomFrac <- biom / sum(biom)
biomRank <- sort(biomFrac, decreasing = T)

# % of total energy allocated to reproduction per species
reproFrac <- getRDI(sim@params) / sum(getRDI(sim@params))
reproRank <- sort(reproFrac, decreasing = T)

# Ranking species with the 3 previous indexes
# Native species will have a fixed rank and invasive species will be put around or maybe not
sparams <- sim@params@species_params
colorRank <- (survivalRank + biomFrac + reproFrac)/3
names(colorRank) <- sim@params@species_params$name

# how many natives species
no_native <- dim(params@species_params)[1]
symbol = c(rep("native",no_native), rep("invasive",(dim(sparams)[1]-no_native)))
label = c(sparams$name[1:no_native],rep(NA,(dim(sparams)[1]-no_native)))


plot_dat <- data.frame(species = sparams$name, beta = sparams$beta,
                       sigma = sparams$sigma, w_inf = sparams$w_inf,
                       w_mat = sparams$w_mat, k_vb = sparams$k_vb,
                       ks = sparams$ks, color = colorRank,
                       Origin = symbol, label = label)
# plot_dat <- reshape2::melt(plot_dat, "species")


ggplot(plot_dat)+
    geom_point(aes(x = w_inf, y = w_mat, color = color, shape = Origin), size = 2) +
    scale_x_continuous(trans = "log10") +
    scale_y_continuous(trans = "log10") +
    ggrepel::geom_text_repel(aes(x = w_inf, y = w_mat, label = label), hjust = 0, nudge_x = 0.05)+
    scale_colour_viridis(name = "Survivability") +
       theme(panel.background = element_blank(), 
         panel.border = element_rect(colour = "gray", fill=NA, size=.5),
         text = element_text(size=10),
         panel.grid.minor = element_line(color = "gray"),
         legend.position = "bottom",legend.key = element_rect(fill = "white"))

ggplot(plot_dat)+
    geom_point(aes(x = beta, y = sigma, color = color, shape = Origin), size = 2) +
    scale_x_continuous(trans = "log10") +
    # scale_y_continuous(trans = "log10") +
     ggrepel::geom_text_repel(aes(x = beta, y = sigma, label = label), hjust = 0, nudge_x = 0.05)+
    scale_colour_viridis(name = "Survivability") +
       theme(panel.background = element_blank(), 
         panel.border = element_rect(colour = "gray", fill=NA, size=.5),
         text = element_text(size=10),
         panel.grid.minor = element_line(color = "gray"),
         legend.position = "bottom",legend.key = element_rect(fill = "white"))

ggplot(plot_dat)+
    geom_point(aes(x = ks, y = k_vb, color = color, shape = Origin), size = 2) +
    # scale_x_continuous(trans = "log10") +
    # scale_y_continuous(trans = "log10") +
     ggrepel::geom_text_repel(aes(x = ks, y = k_vb, label = label), hjust = 0, nudge_x = 0.05)+
    scale_colour_viridis(name = "Survivability") +
       theme(panel.background = element_blank(), 
         panel.border = element_rect(colour = "gray", fill=NA, size=.5),
         text = element_text(size=10),
         panel.grid.minor = element_line(color = "gray"),
         legend.position = "bottom",legend.key = element_rect(fill = "white"))


```

as a function

```{r}

# sim <- readRDS("~/RstudioProjects/RDS_kappaExperiments/kappa0.005/sim1.RDS")

survivalDiag <- function(sim)
{
 
# length of survival per species
survivalLength <- sim@params@species_params$ext - sim@params@species_params$pop # negative values means the species was still alive at the end of the sim
names(survivalLength) <- sim@params@species_params$species
survivalRank <- sort(survivalLength) # this sort put all the surviving species first, then the extinct ones (but in inverted order)
survivalRank <- c(survivalRank[which(survivalRank <0)],survivalRank[rev(which(survivalRank >= 0))]) # now they are in the right order, need to assign weights

# for now, you get 1 if you survived and 0 if you didn't
survivalRank[which(survivalRank >= 0)] <- 0
survivalRank[which(survivalRank < 0)] <- 1
survivalRank <- survivalRank[order(as.numeric(names(survivalRank)))]
    
# doing different order because don't know which one I pref

# % of total community biomass per species
biom <- getBiomass(sim)[dim(sim@n)[1],]
biomFrac <- biom / sum(biom)
biomRank <- sort(biomFrac, decreasing = T)

# % of total energy allocated to reproduction per species
reproFrac <- getRDI(sim@params) / sum(getRDI(sim@params))
reproRank <- sort(reproFrac, decreasing = T)

# Ranking species with the 3 previous indexes
# Native species will have a fixed rank and invasive species will be put around or maybe not
sparams <- sim@params@species_params
colorRank <- (survivalRank + biomFrac + reproFrac)/3
names(colorRank) <- sim@params@species_params$name

# how many natives species
no_native <- length(which(sim@n[1,,1]>0)) # species with positive abundance at sim start
symbol = c(rep("native",no_native), rep("invasive",(dim(sparams)[1]-no_native)))
label = c(sparams$name[1:no_native],rep(NA,(dim(sparams)[1]-no_native)))


plot_dat <- data.frame(species = sparams$name, beta = sparams$beta,
                       sigma = sparams$sigma, w_inf = sparams$w_inf,
                       w_mat = sparams$w_mat, k_vb = sparams$k_vb,
                       ks = sparams$ks, R_max = sparams$R_max, color = colorRank,
                       Origin = symbol, label = label)

return(plot_dat)
}




```


### Exploring kappa setup


Let's do a bunch of sims per kappa
10 sims per value, gradient of 10 kappas, makes 100 simulations

```{r}

kappa_vec <- c(0.0005, 0.001, 0.005,0.01, 0.05,0.1, 0.5, 1, 5)

for(ikappa in kappa_vec)
{
  if(!dir.exists(paste0("~/RstudioProjects/RDS_kappaExperimentsLong/kappa",ikappa))) dir.create(paste0("~/RstudioProjects/RDS_kappaExperimentsLong/kappa",ikappa))
          params <- evoParams(no_sp = 9,kappa = ikappa)
     for(iSim in 1:10)
     {
       sim <- evoProject(params = params, mutation = 0, alien = 5, t_max = 3000)
       saveRDS(sim, file = paste0("~/RstudioProjects/RDS_kappaExperimentsLong/kappa",ikappa,"/sim",iSim,".RDS"))
     }
}

```

result analysis


## Trying with no Rmax


```{r}
# 
# 
#           params <- evoParams(no_sp = 9,kappa = 5)
#           params@species_params$R_max <- Inf
# sim <- evoProject(params = params, mutation = 0, alien = 5, t_max = 200)
# plotBiomass(sim)


kappa_vec <- c(0.0005, 0.001, 0.005,0.01, 0.05,0.1, 0.5, 1, 5)

for(ikappa in kappa_vec)
{
  if(!dir.exists(paste0("simulations/noRmax/kappa",ikappa))) dir.create(paste0("simulations/noRmax/kappa",ikappa))
          params <- evoParams(no_sp = 9,kappa = ikappa)
          params@species_params$R_max <- Inf
     for(iSim in 1:3)
     {
       sim <- evoProject(params = params, mutation = 0, alien = 5, t_max = 1500)
       saveRDS(sim, file = paste0("simulations/noRmax/kappa",ikappa,"/sim",iSim,".RDS"))
     }
}

```




need to pool sims per kappa and plot kappa facets


```{r}
tic()
kappa_vec <- c(0.0005, 0.001, 0.005, 0.01, 0.05, 0.1, 0.5, 1, 5) 
folderName <- "simulations/noRmax"
plot_dat <- NULL
for(iFolder in kappa_vec)
{
simFolder <- paste0(folderName,"/kappa",iFolder,"/")

kappa_dat <- NULL

for (iSim in 1:length(dir(simFolder))) {
  
  if(file.exists(paste0(simFolder,"sim",iSim,".RDS"))) {
    
    mySim <- readRDS(paste0(simFolder,"sim",iSim,".RDS"))
    
    res <- survivalDiag(mySim)
    
    kappa_dat <- rbind(res,kappa_dat)
  }
  
}

# all sims are issued from the same param so the native species will be duplicated
# same names but not necessarily same survival score
# need to do an average of them

# native species
nativeName <- unique(kappa_dat$label)
nativeName <- nativeName[!is.na(nativeName)]
native_df <- NULL
for(iSpecies in nativeName)
{
  temp <- filter(kappa_dat, species == iSpecies)
  avgColor <- mean(temp$color)
  
  temp_df <- temp[1,]
  temp$color <- avgColor
  
  native_df <- rbind(native_df,temp_df)
}

kappa_dat2 <- filter(kappa_dat, !(label %in% nativeName))
kappa_dat <- rbind(native_df,kappa_dat2)

kappa_dat$kappa <- iFolder

plot_dat <- rbind(plot_dat, kappa_dat)

}
toc()

saveRDS(plot_dat, file = paste0(folderName,"/survivalScore.rds"))
```


```{r}
p1 <- ggplot(plot_dat)+
    geom_point(aes(x = w_inf, y = w_mat, color = color, shape = Origin, size = color)) +
    scale_x_continuous(trans = "log10") +
    scale_y_continuous(trans = "log10") +
    ggrepel::geom_text_repel(aes(x = w_inf, y = w_mat, label = label), hjust = 0, nudge_x = 0.05)+
  facet_wrap(~kappa) +
    scale_colour_viridis(name = "Survivability") +
       theme(panel.background = element_blank(), 
         panel.border = element_rect(colour = "gray", fill=NA, size=.5),
         text = element_text(size=10),
         panel.grid.minor = element_line(color = "gray"),
         legend.position = "bottom",legend.key = element_rect(fill = "white"))


p2 <- ggplot(plot_dat)+
    geom_point(aes(x = beta, y = sigma, color = color, shape = Origin, size = color)) +
    scale_x_continuous(trans = "log10") +
    # scale_y_continuous(trans = "log10") +
     ggrepel::geom_text_repel(aes(x = beta, y = sigma, label = label), hjust = 0, nudge_x = 0.05)+
    facet_wrap(~kappa) +
    scale_colour_viridis(name = "Survivability") +
       theme(panel.background = element_blank(), 
         panel.border = element_rect(colour = "gray", fill=NA, size=.5),
         text = element_text(size=10),
         panel.grid.minor = element_line(color = "gray"),
         legend.position = "bottom",legend.key = element_rect(fill = "white"))


p3 <- ggplot(plot_dat)+
    geom_point(aes(x = ks, y = k_vb, color = color, shape = Origin, size = color)) +
    # scale_x_continuous(trans = "log10") +
    # scale_y_continuous(trans = "log10") +
     ggrepel::geom_text_repel(aes(x = ks, y = k_vb, label = label), hjust = 0, nudge_x = 0.05)+
    facet_wrap(~kappa) +
    scale_colour_viridis(name = "Survivability") +
       theme(panel.background = element_blank(), 
         panel.border = element_rect(colour = "gray", fill=NA, size=.5),
         text = element_text(size=10),
         panel.grid.minor = element_line(color = "gray"),
         legend.position = "bottom",legend.key = element_rect(fill = "white"))

```


with a lots of points, make it as a heat map or kknn

also review survivability score and gradient

### plotting number of phen


```{r}

# sim <- readRDS("~/RstudioProjects/RDS_kappaExperiments/kappa0.005/sim1.RDS")



plotNoSp <- function(sim, returnData = F)
{
  
  myData <- data.frame(pop = sim@params@species_params$pop, ext = sim@params@species_params$ext)
  t_max = dim(sim@n)[1]
  totalSp <- vector(mode = "numeric", length = t_max)
  totalSp[1] <- length(which(1 == myData$pop))
  totalExt <- vector(mode = "numeric", length = t_max)

  for(iTime in 2:t_max) totalSp[iTime] <- totalSp[iTime-1] + length(which(iTime == myData$pop))
    
  for(iTime in 2:t_max) totalExt[iTime] <- totalExt[iTime-1] + length(which(iTime == myData$ext))
  
  plot_dat <- data.frame(time = 1:t_max, totalSp = totalSp, extinction = totalExt, speciesNumber = totalSp - totalExt)  
  
  
 p <- ggplot(plot_dat)+
    geom_line(aes(x = time, y = speciesNumber))
    
  
 
  if(returnData) return(plot_dat) else return(p)
  
}



```


```{r}



tic()
kappa_vec <- c(0.0005, 0.001, 0.005, 0.01, 0.05, 0.1, 0.5, 1, 5) 
plot_dat <- NULL
for(iFolder in kappa_vec)
{
simFolder <- paste0(folderName,"/kappa",iFolder,"/")

kappa_dat <- NULL

for (iSim in 1:length(dir(simFolder))) {
  
  if(file.exists(paste0(simFolder,"sim",iSim,".RDS"))) {
    
    mySim <- readRDS(paste0(simFolder,"sim",iSim,".RDS"))
    
    res <- plotNoSp(mySim, returnData = T)
    
    kappa_dat <- cbind(res$speciesNumber,kappa_dat)
  }
  
}
kappa_datMean <- apply(kappa_dat,1,mean )

plot_dat <- cbind(plot_dat, kappa_datMean)

}
colnames(plot_dat) <- kappa_vec
toc()

saveRDS(plot_dat, file = paste0(folderName,"/noSp.rds"))
```


## Extracting average total biomass per kappa


```{r}
tic()
kappa_vec <- c(0.0005, 0.001, 0.005, 0.01, 0.05, 0.1, 0.5, 1, 5) 
plot_dat <- NULL
for(iFolder in kappa_vec)
{
simFolder <- paste0(folderName,"/kappa",iFolder,"/")

kappa_dat <- NULL

for (iSim in 1:length(dir(simFolder))) {
  
  if(file.exists(paste0(simFolder,"sim",iSim,".RDS"))) {
    
    mySim <- readRDS(paste0(simFolder,"sim",iSim,".RDS"))
    
    res <- getBiomass(mySim)
    res <- rowSums(res)
    
    kappa_dat <- cbind(kappa_dat,res)
  }
  
}
kappa_datMean <- apply(kappa_dat,1,mean )

plot_dat <- cbind(plot_dat, kappa_datMean)

}
colnames(plot_dat) <- kappa_vec
toc()

saveRDS(plot_dat, file = paste0(folderName,"/totBiom.rds"))
```

